Tight Generalization Bounds for Hard Parameter Sharing in High-dimensional Linear Regression

### 2020-10-09 arxiv/diff_cov.tex
revise

Our key result characterizes the asymptotic limit of the sum of two arbitrarily different sample covariance matrices.
To motivate our study, consider a special case where $A$ is equal to one.
Without loss of generality, we consider task two's prediction loss and the same result applies to task one.
By solving $B$ in equation \eqref{eq_mtl}, we obtain the hard parameter sharing estimator $\hat{\beta}_2^{\MTL} = B A_2 = B$ as follows:
\begin{align*}%\label{def hatsig}
	\hat{\beta}_2^{\MTL} = {\hat{\Sigma}}^{-1} ({X^{(1)}}^{\top} Y^{(1)} + {X^{(2)}}^{\top} Y^{(2)}), \text{ where } \hat{\Sigma} = {X^{(1)}}^{\top} X^{(1)} + {X^{(2)}}^{\top} X^{(2)}.
\end{align*}
The matrix $\hat{\Sigma}$ adds up both tasks' sample covariances, and the expectation of $\hat{\Sigma}$ is equal to a mixture of their population covariance matrices, with mixing proportions determined by their sample sizes.
The empirical loss of $\hat{\beta}_2^{\MTL}$ is close to its expectation similar to Claim \ref{claim_pred_err}.
Hence, we consider the expected loss conditional on the covariates as follows:
\begin{align}
	\exarg{\cE}{L(\hat{\beta}_2^{\MTL}) \mid X^{(1)}, X^{(2)}} =
	& \bignorm{(\Sigma^{(2)})^{1/2} \hat{\Sigma}^{-1} {X^{(1)}}^{\top} X^{(1)} (\beta^{(1)} - \beta^{(2)})}^2 \label{eq_bias_2task} \\
	& + \sigma^2 \bigtr{\Sigma^{(2)}\hat{\Sigma}^{-1}}. \label{eq_variance_2task}
\end{align}

>>>>

Our key result characterizes the asymptotic limit of the sum of two arbitrarily different sample covariance matrices.
To motivate our study, consider a special case where $A$ is equal to one.
Without loss of generality, we consider task two's prediction loss and the same result applies to task one.
By solving $B$ in equation \eqref{eq_mtl}, we obtain the hard parameter sharing estimator $\hat{\beta}_2^{\MTL} = B A_2 = B$ as follows:
\begin{align}\label{def hatsig}
	\hat{\beta}_2^{\MTL} = {\hat{\Sigma}}^{-1} ({X^{(1)}}^{\top} Y^{(1)} + {X^{(2)}}^{\top} Y^{(2)}), \text{ where } \hat{\Sigma} = {X^{(1)}}^{\top} X^{(1)} + {X^{(2)}}^{\top} X^{(2)}.
\end{align}
The matrix $\hat{\Sigma}$ adds up both tasks' sample covariances, and the expectation of $\hat{\Sigma}$ is equal to a mixture of their population covariance matrices, with mixing proportions determined by their sample sizes.
The empirical loss of $\hat{\beta}_2^{\MTL}$ is close to its expectation similar to Claim \ref{claim_pred_err}.
Hence, we consider the expected loss conditional on the covariates as follows:
\begin{align}
	\exarg{\cE}{L(\hat{\beta}_2^{\MTL}) \mid X^{(1)}, X^{(2)}} =
	& \bignorm{(\Sigma^{(2)})^{1/2} \hat{\Sigma}^{-1} {X^{(1)}}^{\top} X^{(1)} (\beta^{(1)} - \beta^{(2)})}^2 \label{eq_bias_2task} \\
	& + \sigma^2 \bigtr{\Sigma^{(2)}\hat{\Sigma}^{-1}}. \label{eq_variance_2task}
\end{align}

### 2020-10-09 arxiv/proof_examples.tex
revise

Using the concentration estimates in Lemma \ref{largedeviation}, we first show that $g(x)$ and its conditional expectation over $X^{(1)}, X^{(2)}$ are close. %$\exarg{\epsilon^{(1)}, \epsilon^{(2)}}{g(x) \mid X_1, X_2 ,\beta_1,\beta_2}$.
 \begin{claim}\label{claim_largedev1}
 In the setting of Corollary \ref{cor_MTL_loss}, we have that with high probability
$$\left| g(x) - \exarg{\epsilon^{(1)}, \epsilon^{(2)} ,\beta_1,\beta_2}{g(x) \mid X_1, X_2 }\right| \le p^{1/2+c} \left(\sigma^2 + \kappa^2+d^2 \right)$$.
%for any small constant $c>0$.
	% \HZ{use consistent $e_1, e_2, e$, etc. Why not just use $\e$? Don't create too many different symbols. It's bad style..}
 \end{claim}
 \begin{proof}
By Fact \ref{lem_minv} (iii), we have that with high probability,
 \be\label{op_X12}
\|X^{(1)}\|\le \sqrt{(\sqrt{n_1} + \sqrt{p})^2 + n_1 \cdot p^{-c_{\varphi}}}\lesssim \sqrt{p}, \quad \|X^{(2)}\|\le \sqrt{(\sqrt{n_2} + \sqrt{p})^2 + n_2 \cdot p^{-c_{\varphi}}}\lesssim \sqrt{p},
 \ee
 and
 \be\label{op_Sig1}
\| \hat\Sigma(x)^{-1}\|\le \frac1{x^2[(\sqrt{n_1} - \sqrt{p})^2 - n_1 \cdot p^{-c_{\varphi}}]+[(\sqrt{n_2} - \sqrt{p})^2 - n_2 \cdot p^{-c_{\varphi}}] }\lesssim \frac{1}{(x^2+1)p},
 \ee
where we used that $1+\tau\le \rho_1=n_1/p\le \tau^{-1}$ and $1+\tau\le \rho_2=n_2/p\le \tau^{-1}$ for a small constant $\tau>0$.

@@

@@

\paragraph{Part $2$: reducing the random-effects model to Theorem \ref{thm_main_RMT}.}
\HZ{restructure this part. merge the result of B.17 and B.18, and state it upfront. separate its proof into two parts}

\HZ{just say that this calculation of B.97 is similar to B.81?}
We provide the predication loss using the optimal $\hat B$ in equation \eqref{eq_Bhat_2task} and optimal $\hat x=\hat x_1/\hat x_2$:
\begin{align}
L(\hat{\beta}_2^{\MTL}) &=\left\|\Sigma_2^{1/2} \left( \hat B \hat x_2 - \beta_2\right)\right\| \nonumber\\
&=  \left\| (\Sigma^{(2)})^{1/2}\hat \Sigma(\hat x)^{-1} \left[(X^{(1)})^\top X^{(1)} (\hat x\beta_1-\hat x^2\beta_2)+ (X^{(2)})^\top \epsilon^{(2)} + \hat x   (X^{(1)})^\top \epsilon^{(1)} \right]\right\|^2.\label{Lbeta_HPS}
\end{align}
%where we denoted $$\hat \Sigma:= \hat x^2 (X^{(1)})^\top X^{(1)}  + (X^{(2)})^\top X^{(2)} .$$
To illustrate the idea of bias-variance decomposition, we assume that $\hat x$ is independent of $\epsilon^{(1)}$, $\epsilon^{(2)}$, $\beta_1$ and $\beta_2$ for now. Then we can calculate the partial expectation of $L(\hat{\beta}_2^{\MTL})$ over the randomness in $\epsilon_1$ and $\epsilon_2$:
 \begin{align}
	\exarg{\epsilon^{(1)}, \epsilon^{(2)}}{L(\hat{\beta}_2^{\MTL}) \mid X_1, X_2,\beta_1,\beta_2}  &=  \bignorm{(\Sigma^{(2)})^{1/2}\hat\Sigma(\hat x)^{-1} (X^{(1)})^{\top}X^{(1)} (\hat x \beta_1 - \hat x^2\beta_2)}^2 \nonumber\\ %\label{eq_bias_2task} 
			& + \sigma^2\cdot \bigtr{ X^{(2)} \hat\Sigma(\hat x)^{-1}\Sigma^{(2)}\hat\Sigma(\hat x)^{-1} (X^{(2)})^\top} \nonumber\\
			&+ \sigma^2\cdot \hat x^2 \bigtr{ X^{(1)} \hat\Sigma(\hat x)^{-1}\Sigma^{(2)}\hat\Sigma(\hat x)^{-1} (X^{(1)})^\top}\nonumber\\
			%+\sigma^2\cdot \bigtr{\Sigma_2(\frac{W_1^2}{W_2^2} X_1^{\top}X_1 + X_2^{\top}X_2)^{-1} } \nonumber\\ 
			&= \bignorm{(\Sigma^{(2)})^{1/2}\hat\Sigma(\hat x)^{-1} (X^{(1)})^{\top}X^{(1)} (\hat x \beta_1 - \hat x^2\beta_2)}^2 +\sigma^2  \tr\left(\Sigma^{(2)}\hat \Sigma(\hat x)^{-1}\right).\nonumber
			%\label{eq_var_2task}
\end{align}
This gives a bias-variance decomposition of the MTL prediction loss: the first term on the right-hand side is the bias term, while the second term is the variance term. For the random-effects model, we further take the partial expectation over $\beta_1$ and $\beta_2$: 
 \begin{align}
	& \exarg{\epsilon^{(1)}, \epsilon^{(2)},\beta_1,\beta_2}{L(\hat{\beta}_2^{\MTL}) \mid X_1, X_2} \nonumber\\
	&	=  \hat x^2 \left[(\hat x-1)^2 \kappa^2 + \frac{(\hat x^2+1)d^2}2\right] \cdot p^{-1}\bigtr{ (X^{(1)})^{\top}X^{(1)} \hat\Sigma(\hat x)^{-1} \Sigma^{(2)} \hat\Sigma(\hat x)^{-1} (X^{(1)})^{\top}X^{(1)} }\nonumber \\
	&+\sigma^2 \tr\left(\Sigma^{(2)}\hat \Sigma(\hat x)^{-1}\right).
			\label{eq_var_2task}
\end{align}
One can compare the above expression with the predication loss for STL, that is,
\begin{align}
	\exarg{\epsilon_2}{L(\hat{\beta}_2^{\STL}) \mid X_2} = \sigma^2 \cdot \bigtr{\Sigma^{(2)} (X_2^{\top} X_2)^{-1}}, \label{eq_var_stl}
\end{align}
we observe that while the bias of $\hat{\beta}_2^{\MTL}$ is always larger than that of $\hat{\beta}_2^{\STL}$, which is zero, the variance of $\hat{\beta}_2^{\MTL}$ is always lower than that of $\hat{\beta}_2^{\STL}$.\footnote{To see why this is true, we apply the Woodbury matrix identity over the second term on the right-hand side of equation \eqref{eq_var_2task} and use the fact that for the product of two PSD matrices, its trace is always nonnegative.}
In other words, training both tasks together helps predict the target task by reducing variance while incurring a bias.
Therefore, whether multi-task learning outperforms single-task learning is determined by the bias-variance decomposition!

@@

In general, for $\hat x$ as  the minimizer of $g(x)$, it depends on $\epsilon^{(1)}$, $\epsilon^{(2)}$, $\beta_1$ and $\beta_2$. However, using the concentration estimates in Lemma \ref{largedeviation}, we can still obtain the expression in equation \eqref{eq_var_2task} up to a small error. In particular, our proof will not use Lemma \ref{lem_hat_v}.

%control the difference between $g(x)$ and its partial expectation. 
 \begin{claim}\label{claim_largedev2}
 In the setting of Proposition \ref{cor_MTL_loss}, if $\hat x=\OO(1)$, then we have that with high probability,
$$\left| L(\hat{\beta}_2^{\MTL})- \cal L(\hat x)\right| \le p^{-1/2+c} \left(\sigma^2 +\kappa^2 +d^2 \right),$$
for any small constant $c>0$. Here we introduced the function

@@

	\cal L(\hat x)	:=  &\hat x^2\left[(\hat x-1)^2 \kappa^2 + \frac{(\hat x^2+1)d^2}2\right] \cdot p^{-1}\bigtr{ (X^{(1)})^{\top}X^{(1)} \hat\Sigma(\hat x)^{-1} \Sigma^{(2)} \hat\Sigma(\hat x)^{-1} (X^{(1)})^{\top}X^{(1)} } \\
	& +\sigma^2\cdot \tr\left(\Sigma^{(2)}\hat \Sigma(\hat x)^{-1}\right).

@@

 \end{claim}
 \begin{proof}

@@

In the setting of Proposition \ref{cor_MTL_loss}, we can further simplify $\cal L(\hat x)$ using Claim \ref{lem_hat_v} and $\Sigma^{(1)}=\Sigma^{(2)}=\id_p$.

\begin{claim}
In the setting of Proposition \ref{cor_MTL_loss}, we have that with high probability,
%\be\label{choiceofpara}
%	\frac{pd^2}{\sigma^2} = \OO(1),\quad p^{-1+c_0} \le \frac{\kappa^2}{\sigma^2 }  \le p^{-\e_0-c_0},
%where $\e_0$ is the constant as appeared in equation \eqref{approxvalid}. Then we have
&\left|\cal L(\hat x)- d^2\cdot {p}^{-1}\tr\left[\hat\Sigma^{-2} \left((X^{(1)})^\top X^{(1)} \right)^2\right] -\sigma^2  \tr\left({\hat \Sigma^{-1}  }\right)\right| \nonumber\\
&\lesssim  \frac{d^4 + \sigma^2 d^2}{\kappa^2}+p^{-1/2+2c}\kappa^2 +  p^{-1/4+c} (\sigma^2 +d^2) , \label{simple1}
for any small constant $c>0$, where we recall that $\hat \Sigma=\hat \Sigma(1)$ as defined in \eqref{def hatsig}.
\end{claim}
\begin{proof}
Using Lemma \ref{lem_hat_v} and equations \eqref{op_X12}-\eqref{op_Sig1}, we can obtain that with high probability,
Applying the estimates \eqref{hatw_add1}, \eqref{est111}, \eqref{est222} and \eqref{est333} to the three terms on the right-hand side, we can conclude \eqref{simple1}.

>>>>

First, using the concentration estimates in Lemma \ref{largedeviation}, we show that $g(x)$ and its conditional expectation over $X^{(1)}, X^{(2)}$ are close. %$\exarg{\epsilon^{(1)}, \epsilon^{(2)}}{g(x) \mid X_1, X_2 ,\beta_1,\beta_2}$.
 \begin{claim}\label{claim_largedev1}
 In the setting of Corollary \ref{cor_MTL_loss}, we have that with high probability
$$\left| g(x) - \exarg{\epsilon^{(1)}, \epsilon^{(2)} ,\beta_1,\beta_2}{g(x) \mid X_1, X_2 }\right| \le p^{1/2+c} \left(\sigma^2 + \kappa^2+d^2 \right)$$.
%for any small constant $c>0$.
	% \HZ{use consistent $e_1, e_2, e$, etc. Why not just use $\e$? Don't create too many different symbols. It's bad style..}
 \end{claim}
 \begin{proof}
By Fact \ref{lem_minv} (iii), we have that with high probability,
 \be\label{op_X12}
\|X^{(1)}\|\le \sqrt{(\sqrt{n_1} + \sqrt{p})^2 + n_1 \cdot p^{-c_{\varphi}}}\lesssim \sqrt{p}, \quad \|X^{(2)}\|\le \sqrt{(\sqrt{n_2} + \sqrt{p})^2 + n_2 \cdot p^{-c_{\varphi}}}\lesssim \sqrt{p},
 \ee
 and
 \be\label{op_Sig1}
\| \hat\Sigma(x)^{-1}\|\le \frac1{x^2[(\sqrt{n_1} - \sqrt{p})^2 - n_1 \cdot p^{-c_{\varphi}}]+[(\sqrt{n_2} - \sqrt{p})^2 - n_2 \cdot p^{-c_{\varphi}}] }\lesssim \frac{1}{(x^2+1)p},
 \ee
where we used that $1+\tau\le \rho_1=n_1/p\le \tau^{-1}$ and $1+\tau\le \rho_2=n_2/p\le \tau^{-1}$ for a small constant $\tau>0$.

@@

Next, bu studying the conditional expectation in \eqref{EgA}, we show that $\hat x$ is close to 1.

@@

\paragraph{Part $2$: reducing the random-effects model to Theorem \ref{thm_main_RMT}.}
%\HZ{restructure this part. merge the result of B.17 and B.18, and state it upfront. separate its proof into two parts}
%\HZ{just say that this calculation of B.97 is similar to B.81?}
We provide the predication loss using the optimal $\hat B$ in equation \eqref{eq_Bhat_2task} and optimal $\hat x=\hat x_1/\hat x_2$:
\begin{align}
L(\hat{\beta}_2^{\MTL}) &=\left\|\Sigma_2^{1/2} \left( \hat B \hat x_2 - \beta_2\right)\right\| \nonumber\\
&=  \left\| (\Sigma^{(2)})^{1/2}\hat \Sigma(\hat x)^{-1} \left[(X^{(1)})^\top X^{(1)} (\hat x\beta_1-\hat x^2\beta_2)+ (X^{(2)})^\top \epsilon^{(2)} + \hat x   (X^{(1)})^\top \epsilon^{(1)} \right]\right\|^2.\label{Lbeta_HPS}
\end{align}
%where we denoted $$\hat \Sigma:= \hat x^2 (X^{(1)})^\top X^{(1)}  + (X^{(2)})^\top X^{(2)} .$$
Using Lemma \ref{lem_hat_v} and the concentration estimates in Lemma \ref{largedeviation}, we can simplify $L(\hat{\beta}_2^{\MTL})$ as in the following claim.

\begin{claim}
In the setting of Proposition \ref{cor_MTL_loss}, we have that with high probability,
%\be\label{choiceofpara}
%	\frac{pd^2}{\sigma^2} = \OO(1),\quad p^{-1+c_0} \le \frac{\kappa^2}{\sigma^2 }  \le p^{-\e_0-c_0},
%\ee
%where $\e_0$ is the constant as appeared in equation \eqref{approxvalid}. Then we have
\begin{align}
&\left|L(\hat{\beta}_2^{\MTL}) - d^2\cdot {p}^{-1}\tr\left[\hat\Sigma^{-2} \left((X^{(1)})^\top X^{(1)} \right)^2\right] -\sigma^2  \tr\left({\hat \Sigma^{-1}  }\right)\right| \nonumber\\
&\lesssim  \frac{d^4 + \sigma^2 d^2}{\kappa^2}+p^{-1/2+2c}\kappa^2 +  p^{-1/4+c} (\sigma^2 +d^2) , \label{simple1}
\end{align}
for any small constant $c>0$, where we recall that $\hat \Sigma=\hat \Sigma(1)$ was defined in \eqref{def hatsig}.
\end{claim}
\begin{proof}
 \iffalse
 To illustrate the idea of bias-variance decomposition, we assume that $\hat x$ is independent of $\epsilon^{(1)}$, $\epsilon^{(2)}$, $\beta_1$ and $\beta_2$ for now. Then we can calculate the partial expectation of $L(\hat{\beta}_2^{\MTL})$ over the randomness in $\epsilon_1$ and $\epsilon_2$:
 \begin{align}
	\exarg{\epsilon^{(1)}, \epsilon^{(2)}}{L(\hat{\beta}_2^{\MTL}) \mid X_1, X_2,\beta_1,\beta_2}  &=  \bignorm{(\Sigma^{(2)})^{1/2}\hat\Sigma(\hat x)^{-1} (X^{(1)})^{\top}X^{(1)} (\hat x \beta_1 - \hat x^2\beta_2)}^2 \nonumber\\ %\label{eq_bias_2task} 
			& + \sigma^2\cdot \bigtr{ X^{(2)} \hat\Sigma(\hat x)^{-1}\Sigma^{(2)}\hat\Sigma(\hat x)^{-1} (X^{(2)})^\top} \nonumber\\
			&+ \sigma^2\cdot \hat x^2 \bigtr{ X^{(1)} \hat\Sigma(\hat x)^{-1}\Sigma^{(2)}\hat\Sigma(\hat x)^{-1} (X^{(1)})^\top}\nonumber\\
			%+\sigma^2\cdot \bigtr{\Sigma_2(\frac{W_1^2}{W_2^2} X_1^{\top}X_1 + X_2^{\top}X_2)^{-1} } \nonumber\\ 
			&= \bignorm{(\Sigma^{(2)})^{1/2}\hat\Sigma(\hat x)^{-1} (X^{(1)})^{\top}X^{(1)} (\hat x \beta_1 - \hat x^2\beta_2)}^2 +\sigma^2  \tr\left(\Sigma^{(2)}\hat \Sigma(\hat x)^{-1}\right).\nonumber
			%\label{eq_var_2task}
\end{align}
This gives a bias-variance decomposition of the MTL prediction loss: the first term on the right-hand side is the bias term, while the second term is the variance term. For the random-effects model, we further take the partial expectation over $\beta_1$ and $\beta_2$: 
 \begin{align}
	& \exarg{\epsilon^{(1)}, \epsilon^{(2)},\beta_1,\beta_2}{L(\hat{\beta}_2^{\MTL}) \mid X_1, X_2} \nonumber\\
	&	=  \hat x^2 \left[(\hat x-1)^2 \kappa^2 + \frac{(\hat x^2+1)d^2}2\right] \cdot p^{-1}\bigtr{ (X^{(1)})^{\top}X^{(1)} \hat\Sigma(\hat x)^{-1} \Sigma^{(2)} \hat\Sigma(\hat x)^{-1} (X^{(1)})^{\top}X^{(1)} }\nonumber \\
	&+\sigma^2 \tr\left(\Sigma^{(2)}\hat \Sigma(\hat x)^{-1}\right).
			\label{eq_var_2task}
\end{align}
One can compare the above expression with the predication loss for STL, that is,
\begin{align}
	\exarg{\epsilon_2}{L(\hat{\beta}_2^{\STL}) \mid X_2} = \sigma^2 \cdot \bigtr{\Sigma^{(2)} (X_2^{\top} X_2)^{-1}}, \label{eq_var_stl}
\end{align}
we observe that while the bias of $\hat{\beta}_2^{\MTL}$ is always larger than that of $\hat{\beta}_2^{\STL}$, which is zero, the variance of $\hat{\beta}_2^{\MTL}$ is always lower than that of $\hat{\beta}_2^{\STL}$.\footnote{To see why this is true, we apply the Woodbury matrix identity over the second term on the right-hand side of equation \eqref{eq_var_2task} and use the fact that for the product of two PSD matrices, its trace is always nonnegative.}
In other words, training both tasks together helps predict the target task by reducing variance while incurring a bias.
Therefore, whether multi-task learning outperforms single-task learning is determined by the bias-variance decomposition!

@@

In general, for $\hat x$ as  the minimizer of $g(x)$, it depends on $\epsilon^{(1)}$, $\epsilon^{(2)}$, $\beta_1$ and $\beta_2$. However, using the concentration estimates in Lemma \ref{largedeviation}, we can still obtain the expression in equation \eqref{eq_var_2task} up to a small error. In particular, our proof will not use Lemma \ref{lem_hat_v}.
\fi
Our proof is divided into two steps. First, with the concentration estimates in Lemma \ref{largedeviation}, we can simplify $L(\hat{\beta}_2^{\MTL})$ as follows. With high probability, we have that for any small constant $c>0$,
%control the difference between $g(x)$ and its partial expectation. 
% \begin{claim}\label{claim_largedev2}
 %In the setting of Proposition \ref{cor_MTL_loss}, if $\hat x=\OO(1)$, then we have that 
\be\label{claim_largedev2} \left| L(\hat{\beta}_2^{\MTL})- \cal L(\hat x)\right| \le p^{-1/2+c} \left(\sigma^2 +\kappa^2 +d^2 \right),
\ee
where the function $\cal L(\hat x)$ is defined as
\begin{align*}
	\cal L(\hat x)	:=  &\hat x^2\left[(\hat x-1)^2 \kappa^2 + \frac{(\hat x^2+1)d^2}2\right] \cdot p^{-1}\bigtr{ (X^{(1)})^{\top}X^{(1)} \hat\Sigma(\hat x)^{-1} \Sigma^{(2)} \hat\Sigma(\hat x)^{-1} (X^{(1)})^{\top}X^{(1)} } \\
	& +\sigma^2\cdot \tr\left(\Sigma^{(2)}\hat \Sigma(\hat x)^{-1}\right).
\end{align*}
% \end{claim}
We postpone the proof of equation \eqref{claim_largedev2} until we complete the proof of \eqref{simple1}. Next, we can further simplify $\cal L(\hat x)$ using Claim \ref{lem_hat_v} and $\Sigma^{(1)}=\Sigma^{(2)}=\id_p$.
More precisely, we claim that with high probability,
%\begin{claim}
%In the setting of Proposition \ref{cor_MTL_loss}, we have that with high probability,
%\be\label{choiceofpara}
%	\frac{pd^2}{\sigma^2} = \OO(1),\quad p^{-1+c_0} \le \frac{\kappa^2}{\sigma^2 }  \le p^{-\e_0-c_0},
%\ee
%where $\e_0$ is the constant as appeared in equation \eqref{approxvalid}. Then we have
\begin{align}
&\left|\cal L(\hat x)- d^2\cdot {p}^{-1}\tr\left[\hat\Sigma^{-2} \left((X^{(1)})^\top X^{(1)} \right)^2\right] -\sigma^2  \tr\left({\hat \Sigma^{-1}  }\right)\right| \nonumber\\
&\lesssim  \frac{d^4 + \sigma^2 d^2}{\kappa^2}+p^{-1/2+2c}\kappa^2 +  p^{-1/4+c} (\sigma^2 +d^2) , \label{simple12}
\end{align}
for any small constant $c>0$. Combining \eqref{claim_largedev2} and \eqref{simple12}, we immediately obtain \eqref{simple1}. %, where we recall that $\hat \Sigma=\hat \Sigma(1)$ as defined in \eqref{def hatsig}.
%\end{claim}
%\begin{proof}

It remains to prove \eqref{simple12}. Using Lemma \ref{lem_hat_v} and equations \eqref{op_X12}-\eqref{op_Sig1}, we obtain that with high probability,

@@

@@

Applying the estimates \eqref{hatw_add1}, \eqref{est111}, \eqref{est222} and \eqref{est333} to the three terms on the right-hand side, we can conclude \eqref{simple12}.
%\end{proof}
Now we give the proof of the estimate \eqref{claim_largedev2}.
 \begin{proof}[Proof of equation \eqref{claim_largedev2}]

@@

### 2020-10-09 arxiv/tools.tex
revise

\subsection{Random Variables with Bounded Moments}
The following lemma gives (almost) sharp concentration bounds for linear and quadratic forms of bounded supported random variables.
%Besides the proof of Lemma \ref{prop_entry}, we also use it in several other places of this paper, including the proofs in Section \ref{sec_proof_general} and Section \ref{app_iso_cov}.
%It constitutes the main difference between our proof and the one in \cite{KY2}, where the authors used a large deviation bound for random variables with arbitrarily high moments.

\begin{lemma}[Lemma 3.8 of \cite{EKYY1} and Theorem B.1 of \cite{Delocal}]\label{largedeviation}
Let $(x_i)$, $(y_j)$ be independent families of centered and independent random variables, and $(A_i)$, $(B_{ij})$ be families of deterministic complex numbers. Suppose the entries $x_i$ and $y_j$ have variances at most $1$, and $n^{-1/2}x_i$ and $n^{-1/2}y_j$ satisfy the bounded support condition (\ref{eq_support}) with $q\le n^{-\phi}$ for a small constant $\phi>0$. Then we have the following bounds:
%for any fixed $\xi>0$, the following bounds hold with $\xi$-high probability:
\begin{align*}
\Big| \frac{1}{\sqrt{n}}\sum_i A_i x_i \Big\vert \prec  q \max_{i} \vert A_i \vert+ \frac{1}{\sqrt{n}}\Big(\sum_i |A_i|^2 \Big)^{1/2} , \quad &\Big\vert \frac1n\sum_{i,j} x_i B_{ij} y_j \Big\vert \prec q^2 B_d  + qB_o + \frac{1}{n}\Big(\sum_{i\ne j} |B_{ij}|^2\Big)^{{1}/{2}}, \\
\Big\vert \frac1n\sum_{i} (|x_i|^2-\mathbb E|x_i|^2) B_{ii}  \Big\vert  \prec q B_d   , \quad &\Big\vert \frac1n\sum_{i\ne j} \bar x_i B_{ij} x_j \Big\vert  \prec qB_o + \frac{1}{n}\Big(\sum_{i\ne j} |B_{ij}|^2\Big)^{{1}/{2}} ,
\end{align*}
where $B_d:=\max_{i} |B_{ii} |$ and $B_o:= \max_{i\ne j} |B_{ij}|.$ Moreover, if all the moments of $\sqrt{n}x_i$ and $\sqrt{n}y_j$ exist in the sense of equation \eqref{assmAhigh2}, then we have stronger bounds
\begin{align}
\Big\vert \frac1{\sqrt{n}}\sum_i A_i x_i \Big\vert \prec  \frac{1}{\sqrt{n}}\Big(\sum_i |A_i|^2 \Big)^{1/2} , \quad  & \Big\vert \frac1n\sum_{i,j} x_i B_{ij} y_j \Big\vert \prec  \frac{1}{n}\Big(\sum_{i\ne j} |B_{ij}|^2\Big)^{{1}/{2}} ,\label{eq largedev1} \\
 \Big\vert  \frac1n \sum_{i} (|x_i|^2-\mathbb E|x_i|^2) B_{ii}  \Big\vert  \prec \frac1{n}\Big( \sum_i |B_{ii} |^2\Big)^{1/2}  ,\quad & \Big\vert \frac1n\sum_{i\ne j} \bar x_i B_{ij} x_j \Big\vert  \prec  \frac{1}{n}\Big(\sum_{i\ne j} |B_{ij}|^2\Big)^{{1}/{2}} .\label{eq largedev2}
\end{align}
\end{lemma}

>>>>

\subsection{Random Variables with Bounded Moments}
The following lemma gives (almost) sharp concentration bounds for linear and quadratic forms of bounded supported random variables. Here we recall that the stochastic domination ``$\prec$" was defined in Definition \ref{stoch_domination}.
%Besides the proof of Lemma \ref{prop_entry}, we also use it in several other places of this paper, including the proofs in Section \ref{sec_proof_general} and Section \ref{app_iso_cov}.
%It constitutes the main difference between our proof and the one in \cite{KY2}, where the authors used a large deviation bound for random variables with arbitrarily high moments.

\begin{lemma}[Lemma 3.8 of \cite{EKYY1} and Theorem B.1 of \cite{Delocal}]\label{largedeviation}
Let $(x_i)$, $(y_j)$ be independent families of centered and independent random variables, and $(A_i)$, $(B_{ij})$ be families of deterministic complex numbers. Suppose the entries $x_i$ and $y_j$ have variances at most $1$, and $n^{-1/2}x_i$ and $n^{-1/2}y_j$ satisfy the bounded support condition (\ref{eq_support}) with $q\le n^{-\phi}$ for a small constant $\phi>0$. Then we have the following bounds:
%for any fixed $\xi>0$, the following bounds hold with $\xi$-high probability:
\begin{align*}
\Big| \frac{1}{\sqrt{n}}\sum_i A_i x_i \Big\vert \prec  q \max_{i} \vert A_i \vert+ \frac{1}{\sqrt{n}}\Big(\sum_i |A_i|^2 \Big)^{1/2} , \quad &\Big\vert \frac1n\sum_{i,j} x_i B_{ij} y_j \Big\vert \prec q^2 B_d  + qB_o + \frac{1}{n}\Big(\sum_{i\ne j} |B_{ij}|^2\Big)^{{1}/{2}}, \\
\Big\vert \frac1n\sum_{i} (|x_i|^2-\mathbb E|x_i|^2) B_{ii}  \Big\vert  \prec q B_d   , \quad &\Big\vert \frac1n\sum_{i\ne j} \bar x_i B_{ij} x_j \Big\vert  \prec qB_o + \frac{1}{n}\Big(\sum_{i\ne j} |B_{ij}|^2\Big)^{{1}/{2}} ,
\end{align*}
where $B_d:=\max_{i} |B_{ii} |$ and $B_o:= \max_{i\ne j} |B_{ij}|.$ Moreover, if all the moments of $ x_i$ and $ y_j$ exist in the sense of equation \eqref{assmAhigh2}, then we have stronger bounds
\begin{align}
\Big\vert \frac1{\sqrt{n}}\sum_i A_i x_i \Big\vert \prec  \frac{1}{\sqrt{n}}\Big(\sum_i |A_i|^2 \Big)^{1/2} , \quad  & \Big\vert \frac1n\sum_{i,j} x_i B_{ij} y_j \Big\vert \prec  \frac{1}{n}\Big(\sum_{i\ne j} |B_{ij}|^2\Big)^{{1}/{2}} ,\label{eq largedev1} \\
 \Big\vert  \frac1n \sum_{i} (|x_i|^2-\mathbb E|x_i|^2) B_{ii}  \Big\vert  \prec \frac1{n}\Big( \sum_i |B_{ii} |^2\Big)^{1/2}  ,\quad & \Big\vert \frac1n\sum_{i\ne j} \bar x_i B_{ij} x_j \Big\vert  \prec  \frac{1}{n}\Big(\sum_{i\ne j} |B_{ij}|^2\Big)^{{1}/{2}} .\label{eq largedev2}
\end{align}
\end{lemma}
